A significance test in forward stepwise regression

### 2014-03-13 paper/paper.tex
Glint fixed... again
Changed algorithm2e to algorithmicx

% put your definitions there:
\startlocaldefs

\usepackage{graphics,amssymb,color}
\usepackage{graphicx,subfigure}
\usepackage[boxed]{algorithm2e}

@@

\newcommand{\grad}{\nabla}
\newcommand{\V}{\mathcal{V}}\endlocaldefs
\newcommand{\K}{\mathcal{K}}
\newcommand\tf{\widetilde{f}}
\newcommand{\pen}{\mathcal{P}}

@@

Some classical attempts to address this issue include Monte Carlo
estimation of tables of adjusted $F$-statistic values \citep{mc:ftoenter},
and permutation statistics \citep{permutation:stop}. Aside from the works
this paper is based on, there have been other recent attempts to do
inference after model selection. Most of these make use of subsampling
\citep{meinshausen:buhlmann} or data splitting \citep{wasserman:roeder}.
Our approach allows use of the full data and does not require the
extra computation involved in subsampling.
Before describing the full approach we first introduce notation and
specify our implementation of forward stepwise, which is slightly
different from the most commonly used versions.
(\todo Check R's \textit{step} for the group criterion)

@@

\begin{algorithm}
\DontPrintSemicolon
\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
\KwResult{Active set $A$ of variable groups included in the model}
\SetKwFunction{lsfitResidual}{lsfitResidual}
\caption{\small \it Forward stepwise procedure with groups and weights}
\BlankLine
$A \gets \emptyset$\;
$A^c \gets \{ 1, \ldots, G \}$\;
$r_0 \gets y$\;
\For{$s \gets 1$ \KwTo $steps$}{
  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
  $A \gets A \cup \{ g^* \}$\;
  $A^c \gets A^c \backslash \{ g^* \}$\;
  \For{$g \in A^c$}{
    $X_g \gets P X_g$\;
  }
  $r_s \gets P r_{s-1}$\;
}
\KwRet{$A$}
\label{algo:fs}
\end{algorithm}

In the common case where all groups are single variables this version
coincides with the standard implementations of forward stepwise.
We will now use forward stepwise to refer specifically to
Algorithm~\ref{algo:fs} unless otherwise specified.

@@

\subsection{Better p-values}

Maybe just reference \cite{tests:adaptive}? It might be worthwhile to
leave in the connection with \cite{significance:lasso} (getting our
test statistic by not using the approximation)

@@

\item \todo Add the figures to this section
\item \todo Convert to exposition instead of bulleted list
\item \todo Update to reflect latest work
  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
  \item Convergence to the limiting Exp(1) distribution is too slow
    \[
      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
    \]
    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
  \item The limiting distribution only depends on $T$, but we also observe $M$
  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
    \[
      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
    \]
    we can just use the left hand side
  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
  \item Now our p-value is
    \[
      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
    \]   
\end{itemize}

@@

\subsection{Glinternet for hierarchical interactions}
\label{sec:glint}
With many variables, fitting models with pairwise interactions can drastically increase model complexity. \cite{glint} propose a method called \textsc{glinternet} to reduce the statistical complexity of this problem. The method imposes a strong hierarchical constraint on interactions (similar to that in \cite{bien:hierarchical}) where an interaction term can only be included if both its main effects are also included. They accomplish this by creating a design matrix with main effects alone and also with groups including main effects and their first order interactions. Concretely, the expanded design matrix is
\begin{equation}
\label{eq:glintmat}
\tilde X = \begin{pmatrix} X_1 & \cdots & X_G & X_{1:2} & \cdots & X_{1:G} & X_{2:3} & \cdots & X_{(G-1):G}  \end{pmatrix}
\end{equation}
where $X_{g:h}$ is the submatrix including $X_g$, $X_h$, and all
$p_gp_h$  column multiples between columns in group $g$ and columns
in group $h$. For example, if
\[
X_g = \begin{pmatrix} X_{g1} & X_{g2} \end{pmatrix}, \quad
X_h = \begin{pmatrix} X_{h1} & X_{h2} \end{pmatrix}
\]
are two groups each containing two columns, then
\[
X_{g:h} = \begin{pmatrix} X_g & X_h & X_{g1} * X_{h1} & X_{g1} * X_{h2} & X_{g1} * X_{h1} & X_{g2} * X_{h2} \end{pmatrix}
\]
where * denotes the pointwise product (Hadamard product) of vectors (the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times the $i$th entry of $X_{h1}$). Note that this expanded matrix has $\tilde G = G(G+1)/2$ groups.
%For the rest of this section we drop the tilde and denote the expanded design matrix as $X$.
After constructing the design matrix, they proceed by
fitting a group lasso model to the data. Because interaction terms only
appear in groups along with their main effects, the hierarchy condition
holds for the fitted model.

We can also fit a model by forward stepwise on the same expanded design
matrix. The grouping still guarantees that our fitted model will
satisfy the hierarchy condition. To achieve a small computational speed-up, we first modify Algorithm~\ref{algo:fs} to keep track of which main effects have already been included. This reduces the search among potential variables.
%However, we must first modify our implementation of forward stepwise to prevent the possibility of including main effect groups after the main effects are already included in the model as part of an interaction group. To do this we compute the residual using the whole model in each step, rather than by subtracting the effect of the current group being included. For a set of group indices $A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the design matrix containing all the groups in $A$.
% \begin{algorithm}
% \DontPrintSemicolon
% \caption{\small \it Modification to compute residual from full model}
% \BlankLine
% \CommentSty{\# Everything the same as before, additionally:}\;
% %  $P_A \gets I - X_AX_A^\dagger$\;
% %  $r_s \gets P_A r_{s-1}$\;
% \If{$g^*$ is an interaction group}{
%   $g_1, g_2 \gets$ main effects groups of $g^*$\;
%   A^c \gets A^c \slash \{ g_1, g_2 \}\;
% }%\;
% %$r_s \gets y - \text{lsfit}(X_A, y)$\;
% \BlankLine
% \label{algo:fs:full}
% \end{algorithm}
\begin{algorithm}
\DontPrintSemicolon
\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
\KwResult{Active set $A$ of variable groups included in the model}
\SetKwFunction{lsfitResidual}{lsfitResidual}
\caption{\small \it Forward stepwise for glinternet}
\BlankLine
Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
$A \gets \emptyset$\;
$A^c \gets \{ 1, \ldots, G \}$\;
$r_0 \gets y$\;
\For{$s \gets 1$ \KwTo $steps$}{
  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
  $A \gets A \cup \{ g^* \}$\;
  $A^c \gets A^c \backslash \{ g^* \}$\;
  \For{$g \in A^c$}{
    $X_g \gets P X_g$\;
  }
  $r_s \gets P r_{s-1}$\;
}
\KwRet{$A$}
\label{algo:fsgl}
\end{algorithm}

Testing this method by simulation, we constructed signals which have
$k/3$ main effects nonzero but with no interactions and the remaining
$2k/3$ nonzero main effects are matched to each other to form interactions.
This way each nonzero main effect with any nonzero interactions has
exactly one nonzero interaction.
The design matrices are independent Gaussian with all groups of size 1.
Results are shown in Figure~\ref{fig:glint}. The left
panel shows average power of forward stepwise. Power is calculated
using the group structure we impose, and not in terms of the original
main effects and interactions. However, the dotted line shows a more
forgiving definition of power where we are rewarded for discovering part
of a nonzero group, i.e. for discovering only one main effect from a
true interaction group. In the right panel we see that the global null
result still holds (red vertical lines), and that with a signal the
p-value is small until the signal has been recovered (blue). The
interaction power is the proportion of nonzero interaction groups
that were discovered and is labeled ``Int. Power.''

\begin{figure}
\begin{center}
\subfigure[Forward stepwise on glinternet]{
\label{fig:glint:fwd}
\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gauss_glint_p20_n200_lower2p9_upper3p6.pdf}}
\hspace{-15pt}
\subfigure[P-value with glinternet]{
\label{fig:glint:pval}
\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n200_p40_g40_k6_lower2p5_upper3p9_glint.pdf}}
\caption{\small \it The left panel shows the power of forward stepwise
for various sparsity levels. The right panel shows our p-value marginally
over each step, with red indiciating the global null case and blue
indicating a signal with $k=6$ nonzero groups.}
\label{fig:glint}
\end{center}
\end{figure}

>>>>

% put your definitions there:
\startlocaldefs
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphics,amssymb,color}
\usepackage{graphicx,subfigure}
%\usepackage[boxed]{algorithm2e}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

@@

\newcommand{\grad}{\nabla}
\newcommand{\V}{\mathcal{V}}\endlocaldefs
\newcommand{\K}{\mathcal{K}}
\newcommand\tf{\widetilde{f}}
\newcommand{\pen}{\mathcal{P}}
\newcommand{\gstar}{g^*}

@@

Some classical attempts to address this issue include Monte Carlo
estimation of tables of adjusted $F$-statistic values \citep{mc:ftoenter},
and permutation statistics \citep{permutation:stop}. Aside from the works
this paper is based on, there have been other recent attempts to do
inference after model selection. Most of these make use of subsampling
\citep{meinshausen:buhlmann} or data splitting \citep{wasserman:roeder}.
Our approach allows use of the full data and does not require the
extra computation involved in subsampling.
Before describing the full approach we first introduce notation and
specify our implementation of forward stepwise, which is slightly
different from the most commonly used versions.

@@

% \begin{algorithm}
% \DontPrintSemicolon
% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
% \KwResult{Active set $A$ of variable groups included in the model}
% \SetKwFunction{lsfitResidual}{lsfitResidual}
% \caption{\small \it Forward stepwise procedure with groups and weights}
% \BlankLine
% $A \gets \emptyset$\;
% $A^c \gets \{ 1, \ldots, G \}$\;
% $r_0 \gets y$\;
% \For{$s \gets 1$ \KwTo $steps$}{
%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
%   $A \gets A \cup \{ g^* \}$\;
%   $A^c \gets A^c \backslash \{ g^* \}$\;
%   \For{$g \in A^c$}{
%     $X_g \gets P X_g$\;
%   }
%   $r_s \gets P r_{s-1}$\;
% }
% \KwRet{$A$}
% \label{algo:fs}
% \end{algorithm}

\begin{algorithm}
  \caption{Forward stepwise variant with groups and weights}
  \label{algo:fs}
  \begin{algorithmic}[1]
    \REQUIRE An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ variable groups with weights $w_g$
    \ENSURE Ordered active set $A$ of variable groups included in the model at each step
    \STATE $A \gets \emptyset$
    \STATE $A^c \gets \{ 1, \ldots, G\}$
    \FOR{$s=1$ to $steps$}
    \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$
    \STATE $P \gets I - X_{g^*}X_{g^*}^\dagger$
    \STATE $A \gets A \cup \{ g^* \}$
    \STATE $A^c \gets A^c \backslash \{ g^* \}$
    \FORALL{$g \in A^c$}
      \STATE $X_g \gets P X_g$
    \ENDFOR
    \STATE $r_s \gets P r_{s-1}$
    \ENDFOR
    \RETURN $A$
  \end{algorithmic}
\end{algorithm}

In the common case where all groups are single variables this version
coincides with the standard implementations of forward stepwise.
We will now use forward stepwise to refer specifically to
Algorithm~\ref{algo:fs} unless otherwise specified. Note that other
implementations of forward stepwise use different criteria for choosing
the next variable to add, such as the correlation with the residual.
Since we do not renormalize the columns after projecting the covariates
(lines 8 to 10 above), and since we have weights, we are not computing
correlations unless the design matrix is orthogonal and all weights are 1.
There are advantages and disadvantages to both criteria which we do not
discuss. Our choice was motivated by the group lasso result in
\cite{tests:adaptive}, but we believe other criteria can be handled
with appropriate modifications.

@@

\subsection{Relation to covariance test}

Consider the simple case for the first covariate to enter the model.
In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
\todo Does $M = \lambda_2$? Check proof, now that statistic has changed

This is Lemma 5 from \cite{significance:lasso}.

@@

  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
  \item Convergence to the limiting Exp(1) distribution is too slow
    \[
      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
    \]
    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
  \item The limiting distribution only depends on $T$, but we also observe $M$
  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
    \[
      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
    \]
    we can just use the left hand side
  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
  \item Now our p-value is
    \[
      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
    \]   
\end{itemize}

\subsection{Exact p-value calculation}

We now describe how to calculate the p-value, following the discussion of
group lasso in \cite{tests:adaptive}. Let $g = \gstar$ be the index of
the group attaining the maximum on line 4 of Algorithm~\ref{algo:fs},
and let $r = \text{rank}(X_g)$ be the rank of the group attaining the maximum.

@@

\subsection{Glinternet for hierarchical interactions}
\label{sec:glint}
In regression settings with many variables, considering models with pairwise interactions can drastically increase model complexity. \cite{glint} propose a method called \textsc{glinternet} to reduce the statistical complexity of this problem. The method imposes a strong hierarchical constraint on interactions (similar to that in \cite{bien:hierarchical}) where an interaction term can only be included if both its main effects are also included. They accomplish this by creating a design matrix with both main effects alone and also with groups including main effects with their first order interactions. Then they fit a group lasso model with the expanded design matrix. Because interaction terms only appear in groups along with their respective main effects, the hierarchy condition holds for the fitted model. We now consider a related procedure as an example problem, but first modify their method to simplify some parts. Let the expanded design matrix be given by
\begin{equation}
\label{eq:glintmat}
\tilde X = \begin{pmatrix} X_1 & \cdots & X_G & X_{1:2} & \cdots & X_{1:G} & X_{2:3} & \cdots & X_{(G-1):G}  \end{pmatrix}
\end{equation}
where $X_{g:h}$ is the submatrix including $X_g$, $X_h$, and all
$p_gp_h$  column multiples between columns in group $g$ and columns
in group $h$. For example, if
\[
X_g = \begin{pmatrix} X_{g1} & X_{g2} \end{pmatrix}, \quad
X_h = \begin{pmatrix} X_{h1} & X_{h2} \end{pmatrix}
\]
are two groups each containing two columns, then
\[
X_{g:h} = \begin{pmatrix} X_g & X_h & X_{g1} * X_{h1} & X_{g1} * X_{h2} & X_{g1} * X_{h1} & X_{g2} * X_{h2} \end{pmatrix}
\]
where * denotes the pointwise product (Hadamard product) of vectors
(the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times
the $i$th entry of $X_{h1}$). Note that this expanded matrix has
$\tilde G = G + \binom{G}{2} = O(G^2)$ groups. We refer to the first
$G$ of these as main effects or main effect groups, and the remaining
as interaction groups. Finally, instead of fitting a
model by group lasso, we use forward stepwise on the expanded design
matrix. The overlapping groups still guarantee that our fitted model will
satisfy the strong hierarchy condition.

%However, we must first modify our implementation of forward stepwise to prevent the possibility of including main effect groups after the main effects are already included in the model as part of an interaction group. To do this we compute the residual using the whole model in each step, rather than by subtracting the effect of the current group being included. For a set of group indices $A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the design matrix containing all the groups in $A$.
% \begin{algorithm}
% \DontPrintSemicolon
% \caption{\small \it Modification to compute residual from full model}
% \BlankLine
% \CommentSty{\# Everything the same as before, additionally:}\;
% %  $P_A \gets I - X_AX_A^\dagger$\;
% %  $r_s \gets P_A r_{s-1}$\;
% \If{$g^*$ is an interaction group}{
%   $g_1, g_2 \gets$ main effects groups of $g^*$\;
%   A^c \gets A^c \slash \{ g_1, g_2 \}\;
% }%\;
% %$r_s \gets y - \text{lsfit}(X_A, y)$\;
% \BlankLine
% \label{algo:fs:full}
% \end{algorithm}
% \begin{algorithm}
% \DontPrintSemicolon
% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
% \KwResult{Active set $A$ of variable groups included in the model}
% \SetKwFunction{lsfitResidual}{lsfitResidual}
% \caption{\small \it Forward stepwise for glinternet}
% \BlankLine
% Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
% $A \gets \emptyset$\;
% $A^c \gets \{ 1, \ldots, G \}$\;
% $r_0 \gets y$\;
% \For{$s \gets 1$ \KwTo $steps$}{
%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
%   $A \gets A \cup \{ g^* \}$\;
%   $A^c \gets A^c \backslash \{ g^* \}$\;
%   \For{$g \in A^c$}{
%     $X_g \gets P X_g$\;
%   }
%   $r_s \gets P r_{s-1}$\;
% }
% \KwRet{$A$}
% \label{algo:fsgl}
% \end{algorithm}

To demonstrate this method by simulation, we constructed signals which have
the first $k/3$ main effects nonzero but with no interactions, and the
remaining $2k/3$ nonzero main effects are matched to each other to form
interactions.
This way each nonzero main effect with any nonzero interactions has
exactly one nonzero interaction. Furthermore, each nonzero interaction
coefficient has been inflated to be larger than the corresponding main
effect coefficients. This special case is favorable for our algorithm,
but our purpose here is merely to demonstrate the flexibility of the
hypothesis test and not to propose a general procedure for models with
interactions.

Results are shown in Figure~\ref{fig:glint}. The left
panel shows average power of forward stepwise. Power is calculated
using the group structure we impose, and not in terms of the original
main effects and interactions. However, the dotted line shows a more
forgiving definition of power where we are rewarded for discovering part
of a nonzero group, i.e. for discovering only one main effect from a
true interaction group. In the right panel we see that the global null
result still holds (red vertical lines), and that with a signal the
p-value is small until the signal has been recovered (blue). The
interaction power is the proportion of nonzero interaction groups
that were discovered and is labeled ``Int. Power.''

\begin{figure}
\begin{center}
\subfigure[Forward stepwise on glinternet]{
\label{fig:glint:fwd}
\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gauss_glint_p20_n200_lower2p9_upper3p6.pdf}}
\hspace{-15pt}
\subfigure[P-value with glinternet]{
\label{fig:glint:pval}
\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n200_p40_g40_k6_lower2p5_upper3p9_glint.pdf}}
\caption{\small \it The left panel shows the power of forward stepwise
for various sparsity levels. The right panel shows our p-value marginally
over each step, with red indicating the global null case and blue
indicating a signal with $k=6$ nonzero groups.}
\label{fig:glint}
\end{center}
\end{figure}
